\textit{Red Oak}\\ Proactive Controller Generation for Algorithmic Policy

### 2015-05-11 draft.tex
edits 2nd pass

\section*{Implementation}
%that's the hope...

@@

We then generate concrete rules from a collection of relations and a list of symbolic rules. Concrete rules differ from symbolic rule in that they do not depend on relations. Each symbolic rule expands into a list of concrete rules. For each of the tuple in the relation on which the rule depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple. We repeat this operation on the created list of symbolic rules until all dependant relations are taken care off, at which point the rules are considered concrete. Due to the left-to-right order of collection, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.

For example, taking the following symbolic rules: 
  \begin{lstlisting}[mathescape]
src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
  \end{lstlisting}
  and the relation ``conn'' as \lstinline| {(150,10), (200, 9), (8, 150)}|, we start by expanding the first rule to:
\begin{lstlisting}[mathescape]
src $\in$ ([101, 1000] $\cap$ 150), dst = 10 $\mapsto$ Deliver
src $\in$ ([101, 1000] $\cap$ 200), dst = 9 $\mapsto$ Deliver
src $\in$ ([101, 1000] $\cap$ 8), dst = 150 $\mapsto$ Deliver
 \end{lstlisting}
which reduces to:
\begin{lstlisting}[mathescape]
src = 150, dst = 10 $\mapsto$ Deliver
src = 200, dst = 9 $\mapsto$ Deliver
 \end{lstlisting}
Note the omission of the third concrete rule, resulting from \lstinline|ipsrc| having an empty range. The second symbolic rule is then taken directly as concrete rule, with the packet in the relation having been matched by the previously generated concrete rule, resulting in:
\begin{lstlisting}[mathescape]
src = 150, dst = 10 $\mapsto$ Deliver
src = 200, dst = 9 $\mapsto$ Deliver
src $\in$ [101, 1000] $\mapsto$ Drop
\end{lstlisting}

@@

\lstinline|run| starts by building the symbolic rules for the given policy, and generating the concrete rules assuming all relations to be empty. In sequence, packet are matched with the current concrete rules, recording the forwarding decision. If a packet is sent to the controller, we evaluate the policy on the packet, potentially adding or removing tuples from relations. We then re-generate the concrete rules with the updated relation, before recording the forwarding decision.

@@

\subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
The rules generated by the presented technique ensures that the only packets being forwarded to the controller are those reaching an add or a remove node and thus potentially modify the state of the controller. However, these packets do not always modify the state of the controller. For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent. By using both local and global transformations on the decision tree, we can get closer to an optimal amount of packet forwarded to the controller, where only packets modifying relations are seen by the controller.

>>>>

\section*{Implementation}

We now describe our prototype implementation of \textit{Red Oak} as a library embedded in the Ocaml programming language.

@@

We then generate concrete rules from a collection of relations and a list of symbolic rules. Concrete rules differ from symbolic rules in that they do not depend on relations. Each symbolic rule expands into a list of concrete rules. For each tuple in the relation on which the symbolic rules depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple. We repeat this operation on the created list of symbolic rules until all dependant relations are taken care of, at which point the rules are considered concrete. Due to the prioritization of the resulting rules, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.

For example, taking the following symbolic rules from our earlier stateful firewall example: 
  \begin{lstlisting}[mathescape]
src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
  \end{lstlisting}
The first rules says: "external hosts that are trusted should be let through", and the second rules says: "traffic from untrusted external hosts should be dropped".

  Assume that the trusted connection relation ``conn'' contains the pairs 
  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external src 150 can talk to internal dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
\begin{lstlisting}[mathescape]
src $\in$ ([101, 1000] $\cap$ 150), dst = 10 $\mapsto$ Deliver
src $\in$ ([101, 1000] $\cap$ 200), dst = 9 $\mapsto$ Deliver
src $\in$ ([101, 1000] $\cap$ 8), dst = 150 $\mapsto$ Deliver
 \end{lstlisting}
To satisfy both symbolic constraints, we take the intersection of the constraints, which corresponds to intersection of the constraints on the ranges of each field. These rules reduce to:
\begin{lstlisting}[mathescape]
src = 150, dst = 10 $\mapsto$ Deliver
src = 200, dst = 9 $\mapsto$ Deliver
 \end{lstlisting}
Note the omission of the third concrete rule, resulting from \lstinline|ipsrc| having an empty range. We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules. The final concrete forwarding rules are now:
\begin{lstlisting}[mathescape]
src = 150, dst = 10 $\mapsto$ Deliver
src = 200, dst = 9 $\mapsto$ Deliver
src $\in$ [101, 1000] $\mapsto$ Drop
\end{lstlisting}

@@

\lstinline|run| starts by building the symbolic rules for the given policy, and generating the concrete rules assuming all relations to be empty. In sequence, packets are matched with the current concrete rules, recording the forwarding decision. If a packet is sent to the controller, we evaluate the policy on the packet, potentially adding or removing tuples from relations. We then re-generate the concrete rules with the updated relation, before recording the forwarding decision.

@@

\subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
The rules generated by the presented technique ensures that the only packets being forwarded to the controller are those reaching an add or a remove node and thus potentially modify the state of the controller. However, these packets do not always modify the state of the controller. For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent. By using both local and global transformations on the decision tree, we can get closer to an optimal number of packet forwarded to the controller, where only packets having a real, meaningful effect are seen by the controller.
